---
title: "Kill-off Profile Analysis using payne_profile()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Kill-off Profile Analysis using payne_profile()}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(PayneProfileR)
```

## Introduction

This vignette demonstrates how to use the `payne_profile()` function to analyze caprine dental wear data and generate kill-off profiles following Payne (1973), with options to visualize the Helmer et al. (2007) barplot style and ideal survival models.

## Example Data

```{r}
data <- data.frame(
  stage = c("A", "AB", "B", "C", "D", "E", "F", "CD", "DEF", "GHI", "EFG", "G"),
  count = c(5, 4, 6, 3, 2, 3, 6, 5, 8, 4, 1, 4)
)
```

## Standard Kill-off Profile

```{r}
res <- payne_profile(data)
res$table
res$barplot
res$curve
```

## Helmer-style Barplot

```{r}
res_helmer <- payne_profile(data, ef_merge = TRUE)
res_helmer$barplot_helmer
```

## Survival Curve with Ideal Strategies

```{r}
res_ideal <- payne_profile(data, add_ideal_curves = TRUE)
res_ideal$ideal_curve
```

## References

- Payne, S. (1973). Kill-off patterns in sheep and goats: the mandibles from Asvan Kale. Anatolian Studies, 23, 281–303.
- Helmer, D., Gourichon, L., & Vila, E. (2007). The development of the exploitation of products from Capra and Ovis... Anthropozoologica, 42(2), 41–69.
